Robust co-occurrence quantification for low-dimensional lexical distributional semantics

### 2016-02-27 figures/interaction-cds.tex
Lexical tasks
Figures
Update

\begin{figure*}
  \centering
  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/simlex-ppmi-cds.pdf}
    \caption{\textbf{Context distribution smoothing.}}
    \label{fig:simlex-ppmi-interaction-cds}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{supplement/figures/simlex-ppmi-neg.pdf}
    \caption{\textbf{Shifting.}}
    \label{fig:simlex-ppmi-interaction-neg}
  \end{subfigure}

  \caption{\textbf{SPMI.}}
  \label{fig:simlex-ppmi-interaction}
\end{figure*}

>>>>

\begin{figure*}
  \centering
    \centerline{\includegraphics[width=1.11\textwidth]{supplement/figures/SimLex999-interaction-cds}}
    \caption{\textbf{The effect of context probability estimation and smoothing on SimLex-999.}}
    \label{fig:interaction-cds}
\end{figure*}

### 2016-02-27 paper.tex
Update

%\title{Robust co-occurrence quantification for lexical distributional semantics experiments}
% MP: how about:
\title{Robust co-occurrence quantification for low-dimensional lexical distributional semantics}

@@

In contrast, much work on \emph{compositional} distributional semantics employs vectors with much fewer dimensions: e.g.~\textbf{2K} \cite{Grefenstette:2011:ESC:2145432.2145580,kartsadrqpl2014,milajevs-EtAl:2014:EMNLP2014}, \textbf{3K} \cite{Dinu:2010:MDS:1870658.1870771,milajevs-purver:2014:CVSC} or \textbf{10K} \cite{polajnar-clark:2014:EACL,Baroni2010nouns}. \textbf{MP: say why}

Such a mismatch in vector dimensionality between lexical and compositional tasks gives rise to a number of questions:
\begin{itemize}
\item To what extent does model performance depend on vector dimensionality?
\item Do parameters influence performance of high-dimensional and low-dimensional models in the same way? Can the findings of \newcite{TACL570} be directly applied to low-dimensional models?
\item If not, can we derive suitable parameter selection heuristics which take account of dimensionality?
\end{itemize}

@@

An alternative solution of avoiding infinity is to increment the probability ratio by one:
%
\begin{equation}
  \label{eq:cpmi}
  \operatorname{CPMI}(x, y) = \log\Big( 1 + \frac{P(x,y)}{P(x)P(y)} \Big)
\end{equation}

@@

Many approaches use only \emph{positive} PMI values, as  negative PMI values may not positively contribute to model performance \cite{Turney:2010:FMV:1861751.1861756}. This can be generalised as an additional hyperparameter $k$ (\texttt{neg}) following \newcite{TACL570}, giving our third PMI variant (abbreviated as SPMI):
%
\begin{equation}
  \label{eq:ppmi}
  \operatorname{SPMI_k} = \max (0, \operatorname{PMI}(x, y) - \log k)
\end{equation}
%
We can apply the same idea to CPMI:
%
\begin{equation}
  \label{eq:pcpmi}
  \operatorname{SCPMI_k} = \max (0, \operatorname{CPMI}(x, y) - \log 2k)
\end{equation}

@@

\subsection{Similarity metric}
\label{sec:similarity-metric}

We test a widely used cosine based similarity, as well as correlation \cite{kiela-clark:2014:CVSC}. \textbf{MP: I would move this to next section: it's not really a parameter here}

@@

Table~\ref{tab:parameters} lists parameters used in this work and their values. We will derive the parameter selection heuristics by greedily selecting parameters (\texttt{cds}, \texttt{neg} and similarity measure) that lead to the highest average performance for each frequency, discriminativeness and context combination. Figures~\ref{fig:interaction-cds} and \ref{fig:interaction-neg} show such interaction of \texttt{cds} and \texttt{neg} with other parameters, we don not report similarity behavior due to the limited space.

@@

\subsection{Evaluation of heuristics}
\label{sec:heuristic-evaluation}

>>>>

\title{Robust co-occurrence quantification for low-dimensional lexical distributional semantics}

@@

In contrast, much work on \emph{compositional} distributional semantics employs vectors with much fewer dimensions: e.g.~\textbf{2K} \cite{Grefenstette:2011:ESC:2145432.2145580,kartsadrqpl2014,milajevs-EtAl:2014:EMNLP2014}, \textbf{3K} \cite{Dinu:2010:MDS:1870658.1870771,milajevs-purver:2014:CVSC} or \textbf{10K} \cite{polajnar-clark:2014:EACL,Baroni2010nouns}. These models  assign tensors to functional words: intransitive verbs and adjectives have  tensors of rank 2,  transitive verbs  rank 3, prepositions rank 7. Tensors are highly dimensional objects, for $V$ a vector space with dimension $k$,  a tensor $V \otimes V \cdots \otimes V$ of rank $n$ over it has dimension  $n^k$. In these models, if one takes $n= \textbf{30k}$, then one will end up with   spaces with intractable dimensions  $\textbf{30k}^2, \textbf{30k}^3$ and $\textbf{30k}^7$.

Such a mismatch in vector dimensionality between lexical and compositional tasks gives rise to a number of questions:
\begin{compactitem}
\item To what extent does model performance depend on vector dimensionality?
\item Do parameters influence performance of high-dimensional and low-dimensional models in the same way? Can the findings of \newcite{TACL570} be directly applied to low-dimensional models?
\item If not, can we derive suitable parameter selection heuristics which take account of dimensionality?
\end{compactitem}

@@

An alternative solution of avoiding infinity is to increment the probability ratio by one, we refer to it as \textit{compressed} PMI:
%
\begin{equation}
  \label{eq:cpmi}
  \operatorname{CPMI}(x, y) = \log\Big( 1 + \frac{P(x,y)}{P(x)P(y)} \Big)
\end{equation}

@@

Many approaches use only \emph{positive} PMI values, as  negative PMI values may not positively contribute to model performance \cite{Turney:2010:FMV:1861751.1861756}. This can be generalised as an additional parameter $k$ (\texttt{neg}) following \newcite{TACL570}, giving our third PMI variant (abbreviated as SPMI):
%
\begin{equation}
  \label{eq:ppmi}
  \operatorname{SPMI_k} = \max (0, \operatorname{PMI}(x, y) - \log k)
\end{equation}
%
We can apply the same idea to CPMI:
%
\begin{equation}
  \label{eq:pcpmi}
  \operatorname{SCPMI_k} = \max (0, \operatorname{CPMI}(x, y) - \log 2k)
\end{equation}

@@

@@

Table~\ref{tab:parameters} lists parameters used in this work and their values. We will derive the parameter selection heuristics by greedily selecting parameters (\texttt{cds} and \texttt{neg}) that lead to the highest average performance for each frequency, discriminativeness and context combination. Figures~\ref{fig:interaction-cds} and \ref{fig:interaction-neg} show such interaction of \texttt{cds} and \texttt{neg} with other parameters. We also vary the similarity measure (cosine and correlation  \cite{kiela-clark:2014:CVSC}), but do not report their behaviour in details due to the limited space.

@@

\paragraph{Evaluation of heuristics}
\label{sec:heuristic-evaluation}
